A standardized framework for risk-based assessment of treatment effect heterogeneity in observational healthcare databases

### 2021-03-18 submission/manuscript.rmd
set line width for readability

\newpage
\doublespacing
\linenumbers
# Introduction
Interest in understanding how a treatment’s effect varies across patients—a concept described as heterogeneity of treatment effects (HTE)—has been growing. This concept is central to the agenda for both personalized (or precision) medicine and comparative effectiveness research. More formally, HTE has been defined as non-random variability in the direction or magnitude of a treatment effect, in which the effect is measured using clinical outcomes. Usually, analyses focus on the relative scale, where treatment effects are assessed one at a time in patient subgroups defined from single covariates, an approach that suffers from low power and multiplicity issues. However, even with well-established constant relative effects, treatment benefit (or harm) may vary substantially on the absolute scale.

@@

* IBM MarketScan Medicare Supplemental Beneficiaries (MDCR): Represents health services of retirees (aged 65 or older) in the United States with primary or Medicare supplemental coverage through privately insured fee-for-service, point-of-service or capitated health plans. These data include adjudicated health insurance claims (e.g. inpatient, outpatient and outpatient pharmacy). Additionally, it captures laboratory tests for a subset of the covered lives.
* IBM MarketScan Medicaid (MDCD): Adjudicated US health insurance claims for Medicaid enrollees from multiple states. It includes hospital discharge diagnoses, outpatient diagnoses and procedures and outpatient pharmacy claims as well as ethnicity and Medicare eligibility.
* IBM MarketScan Commercial Claims and Encounters (CCAE): Data from individuals enrolled in US employer-sponsored   insurance   health   plans. The data includes adjudicated health insurance   claims (e.g. inpatient, outpatient, and outpatient pharmacy) as well as enrollment data from large employers and health plans who provide private healthcare coverage to employees, their spouses and dependents. Additionally, it captures laboratory tests for a subset of the covered lives.

>>>>

\newpage 
\doublespacing 
\linenumbers

# Introduction
Interest in understanding how a treatment’s effect varies across patients—a
concept described as heterogeneity of treatment effects (HTE)—has been growing.
This concept is central to the agenda for both personalized (or precision)
medicine and comparative effectiveness research. More formally, HTE has been
defined as non-random variability in the direction or magnitude of a treatment
effect, in which the effect is measured using clinical outcomes. Usually,
analyses focus on the relative scale, where treatment effects are assessed one
at a time in patient subgroups defined from single covariates, an approach that
suffers from low power and multiplicity issues. However, even with
well-established constant relative effects, treatment benefit (or harm) may vary
substantially on the absolute scale.

@@

* IBM MarketScan Medicare Supplemental Beneficiaries (MDCR): Represents health
services of retirees (aged 65 or older) in the United States with primary or
Medicare supplemental coverage through privately insured fee-for-service,
point-of-service or capitated health plans. These data include adjudicated
health insurance claims (e.g. inpatient, outpatient and outpatient pharmacy).
Additionally, it captures laboratory tests for a subset of the covered lives.
* IBM MarketScan Medicaid (MDCD): Adjudicated US health insurance claims for
Medicaid enrollees from multiple states. It includes hospital discharge
diagnoses, outpatient diagnoses and procedures and outpatient pharmacy claims
as well as ethnicity and Medicare eligibility.  IBM MarketScan Commercial
* Claims and Encounters (CCAE): Data from individuals enrolled in US
employer-sponsored   insurance   health   plans. The data includes adjudicated
health insurance   claims (e.g. inpatient, outpatient, and outpatient
pharmacy) as well as enrollment data from large employers and health plans who
provide private healthcare coverage to employees, their spouses and
dependents. Additionally, it captures laboratory tests for a subset of the
covered lives.
